---
fontsize: 11pt
geometry: margin=0.50in
subparagraph: yes
title: St. Geme transposon library mapping project
date: 'October 2019, draft 1'
author: John K. Everett, Ph.D.
indent: false
header-includes:
   - \usepackage{pdflscape}
   - \newcommand{\blandscape}{\begin{landscape}}
   - \newcommand{\elandscape}{\end{landscape}}
   - \usepackage{pdfpages}
output:
  pdf_document:
    toc: false
    keep_tex: no
    latex_engine: xelatex
---
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(RColorBrewer)
library(tidyverse)
load('data/analysis.RData')
ppnum <- function(n) format(n, big.mark = ",", scientific = FALSE, trim = TRUE)
```

This analysis describes the creation of a sequencing library created from *Kingella kingae* DNA samples provided 
by the St. Geme research group and the subsequent mapping of identified transpon insertions. A sequencing library 
was created by shearing genomic DNA and the attachment of adapter sequences followed by a nested PCR 
where the first set of primers bound within the body of the experimental transposon while the second set 
of primers bound within the transposon ITR segments. The library was sequenced with the Illumina MiSeq 
platform and transposon insertions were identified by searching for the 8 terminal ITR nucleotides followed 
by a TA sequence (CAACCTGTTA). The number of insertions recovered from each sample is shown in Table 1. 
Sequences were aligned to the representative *Kingella kingae* strain *Vir5453* (NCBI tax id: 1305785).

For the purpose of visualizing the data, the number of recovered insertions were normalized by dividing the 
number of sites within 10KB genomic blocks by the total number of sites recovered in each sample (Figure 1).  

\vspace{0.5cm}

\small
Figure 1. Visualization of recovered insertions within *Kingella kingae*.  
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=4.3}
genomeInsertionPlot + theme(text = element_text(size=11))
```

\vspace{0.50cm}

\small
Table 1. Number of recovered insertions. 
\normalsize

```{r, echo=FALSE}
names(sites.tbl) <- c('Sample', 'Insertions')
sites.tbl$Insertions <- ppnum(sites.tbl$Insertions)
tab <- bind_cols(sites.tbl[grepl('Control', sites.tbl$Sample),], sites.tbl[grepl('Challenge', sites.tbl$Sample),])
names(tab) <- c('Sample', 'Insertions', 'Sample', 'Insertions')
kable(tab, "latex", longtable = T, linesep = "", booktabs = T) %>% kable_styling(position = "center", font_size = 10)
```

\newpage

The number of insertions within transcription units (TUs) was gauged using two approaches. The first 
approach considered the number of insertions within each TU divided by the total number of insertions 
recovered in the sample. The second approach considered the total number of inferred cells 
(unique genomic break points) associated with insertions within each TU divided by the total number 
of inferred cells in the sample. The site count approach showed a fair degree of variation between techinical 
replicates (Figure 2) while the abundance method showed less variation between replicates and averaged samples (Figure 3).  

\vspace{0.5cm}

\small
Figure 2. Distriubtions of differences between technical replicate and averaged sample TUs using the site count approach. 
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d1.result$diffPlot
```

\vspace{1.2cm}

\small
Figure 3. Distriubtions of differences between technical replicate and averaged sample TUs using the abundance approach. 
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d2.result$diffPlot + labs(y = 'Normalized abundance counts', x = 'Experiment')
```

\newpage

Using the abundance approach, clear clustering of biological samples was found though there was not remarkable separation 
between control and challenge samples whithin biological sample clusters (Figures 4 & 5). The normalized site count 
approach provided less distinctive clustering (Supp. Figures S1 & S2).

\vspace{0.5cm}

\small
Figure 4. Principle component analysis of all samples using the abundance normalization approach.
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d2.result$PCA
```

\vspace{1.2cm}

\small
Figure 5. Principle component analysis of averaged technical replicates using the abundance normalization approach.
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d2.result$PCA_avg
```

\newpage

For each transcription unit, t-tests were used to test for differences between control and challenge insertion frequencies. 
Transcription units with significant uncorrected p-values are shown in Tables 2 & 3. Full gene tables 
are available on-line via this [**link**](https://github.com/everettJK/TransposonMapping_StGeme).

\vspace{1.0cm}

\small
Table 2. Genes with significant uncorrected p-values using the abundance correction method.
\normalsize

```{r, echo = FALSE}
d2.result$enrichmentTable[which(d2.result$enrichmentTable$nearestFeature == 'GeneID:34399865'),]$geneDesc <-  'glutamate–ammonia ligase]-adenylyl-L-tyrosine phosphorylase...'

kable(subset(select(d2.result$enrichmentTable, nearestFeature, geneDesc,  pVal, pVal.adj, higherInChallenge),  pVal <= 0.05), "latex", longtable = T, linesep = "", booktabs = T) %>% kable_styling(position = "left", font_size = 8)
```

\vspace{1.0cm}

\small
Table 3. Genes with significant uncorrected p-values using the normalized site count method.
\normalsize

```{r, echo = FALSE}
kable(subset(select(d1.result$enrichmentTable, nearestFeature, geneDesc,  pVal, pVal.adj,higherInChallenge),  pVal <= 0.05), "latex", longtable = T, linesep = "", booktabs = T) %>% kable_styling(position = "left", font_size = 8)
```

\newpage

**Supplimental**   


\small
Figure S1. Principle component analysis of all samples using the normalized site count method.
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d1.result$PCA
```

\vspace{1.2cm}

\small
Figure S2. Principle component analysis of averaged technical replicates using the normalized site count method.
\normalsize

\vspace{0.10cm}

```{r, echo=FALSE, fig.height=3.7}
d1.result$PCA_avg



